### 2014-07-08 tex/main-body-text.tex
Edits to conclusions.

Despite the similar allocation of individuals into $K$ clusters, the 
plateau analysis to estimate $K$ in {\sc flock} leads to different conclusions among groups of runs, as well as 
one that differs from {\sc structure}. While no plateaus were observed in any of the six groups 
for a $K\geq 2$, the plateau record varied greatly among each
group for $K = 2$ (plateau sequences: Run1 = 3,2,3,7,4; Run2 = 2,5,2,3,2,3; Run3 = 2,2,2,2,2,5; 
Run4 = 2,3,4,2,4,3; Run5 = 2,2,5,5,3; and; Run6 = 2,2,2,2,2,3,3). The plateau analysis 
procedure suggested by \citet{Duc&Tur2012} leads to
either an estimate of K $\geq$ 2 (Run1) or {\em undecided} as a consequence of no plateau 
lengths being greater than 6. The ($\ln P(D)$) from {\sc structure}
was largest at $K$ = 6 which was further supported by the $q_i$ plots, however,
the $\Delta K$ method supported a $K$ of 2.

\hrule

@@

\section*{Conclusions}

One of the major claimed advantages of {\sc flock} over MCMC methods for solving the 
K-partition problem is that, ``The very short processing times for each run of {\sc flock} allows for
 comparing partitions from many runs of each $K$'' \citep[][p.~735]{Duc&Tur2012}.
While the authors observed improved process time in their 
comparisions we did not observe the same advantage in our analysis. In their comparison, the 
run parameters for {\sc flock} were 50 runs and 20 iterations for each $K$ from 1 to four more than 
the true value of $K$. 
These process times were compared to the admixture model with correlated allele 
frequencies in {\sc structure} with ten iterations of a 50,000 sweep burn-in followed 
by a 200,000 sweep sample from the posterior of $K$ from 1 to K+3. 
It seemed reasonable to compare, for each $K$, one iteration of {\sc structure} 
to one iteration of {\sc flock} (50 different starting allocations that are reallocated 20 times) 
as both evaluate the partition space once. The dataset used in this comparision was an order of magnatude
larger than the largest used by \citep{Duc&Tur2012} (N = 60, 120, and 240). Any process speed advantages 
may only been observed with small sample sizes.

The partition space over which {\sc flock} evaluates how individuals are clustered 
greatly increases with both the number of individuals and the number of clusters into 
which they can be partitioned. The total partition space can be 
defined by $\sum\limits_{i=1}^k S(n,k)$ where S(n,k) is a Stirling number of the 
second kind, $k$ is the number of clusters, and $n$ is the number of individuals.
The likelihood surface spanning all possible partitions
may be characterized by multiple peaks and troughs.
The peak or plateau that the  {\sc flock} algorithm finds is entirely dependent on the initial allocation of 
individuals as it has no way to visit areas of the likelihood surface with smaller values.
 It is not surprising that with 2,596 individuals we could have 
drastically different plateau sequences with just two clusters and that as $K$ increases
{\sc flock} finds suboptimal allocations. If the number of initial allocations does not increase
as $K$ is increased, the starting allocations become diffusely scattered in the space of all
possible partitions making it unlikely that they will converge.

For those that do decide to use {\sc flock} as an alternative to {\sc structure}, the adhoc rules 
to estimate K may be overly stringent. While there has been ample 
discourse concering choosing the optimal K in {\sc structure} \citep{Evannoetal2005, 
Pritchardetal2000, Wap&Gag2006} and \citet{Pritchardetal2000} urges users 
to be cautious of interpreting K, much less effort has been put forth examining how the number of individuals
and thenumber of starting allocations can affect not only the convergence of different runs, but also the 
inference of $K$. The authors developed estimation rules based on simulated data 
sets with a modest number of individuals and different migration regimes and rates, number of loci, and K. 
By relying on identical allocations (mean LLOD values) as a means of supporting 
a particular $K$,  similar but not idential 
allocations will provide no support to a given $K$.

>>>>

Despite the similar allocation of individuals into $K$ clusters, the 
plateau analysis to estimate $K$ in {\sc flock} leads to different conclusions among groups of runs, as well as 
one that differs from {\sc structure}. While no plateaus were observed in any of the six groups 
for a $K\geq 2$, the plateau record varied greatly among each
group for $K = 2$ (plateau sequences: Run1 = 3,2,3,7,4; Run2 = 2,5,2,3,2,3; Run3 = 2,2,2,2,2,5; 
Run4 = 2,3,4,2,4,3; Run5 = 2,2,5,5,3; and; Run6 = 2,2,2,2,2,3,3). The plateau analysis 
procedure suggested by \citet{Duc&Tur2012} leads to
either an estimate of K $\geq$ 2 (Run1) or {\em undecided} as a consequence of no plateau 
lengths being greater than 6. The ($\ln P(D)$) from {\sc structure}
was largest at $K$ = 6 which was further supported by the $q_i$ plots, however,
the $\Delta K$ method supported a $K$ of 3.

@@

\section*{Conclusions}
In the past decade, the use of genetic markers to identify population structure and to attempt inference of
the number of genetic clusters ($K$) in a collection has dramatically increased in the fields of ecology, evolution, epidemiology, and conservation
genetics. By virtute of this expanding desire to identify $K$ there has been a push to 
develop fast and easy-to-use methods that can accurately infer $K$. We show that {\sc flock}, a program billed as providing
``more accurate allocations and more reliable estimates of $K$ [while running]
much faster than {\sc structure} \citep[][p.~734]{Duc&Tur2012},'' may not, in fact, represent such 
a radical departure from {\sc structure} itself.  In particular, we have shown that the algorithm implemented 
in {\sc flock} is a simulated annealing algorithm for the Bayesian 
MAP estimate from a marginalized and restricted form of the {\sc structure} model.  
Given such similarities between the two methods it is not surprising that
we observed relatively
congruent results between the ``best run'' from {\sc flock} and {\sc structure} using 
a real data set from \textit{O. mykiss}.

Though we observed very similar results
between the two programs at low values of $K$, {\sc flock} had a tendency to provide inconsistent solutions
among its ``best runs'' at higher values, while still finding solutions that were similar to {\sc structure}'s.
This is likely a consequence of the rapidly increasing size of the search space over possible 
partitions as $K$ and $n$, the number of individuals, increase. 
The likelihood surface defined on partitions of this space is likely to be
be characterized by multiple peaks and troughs.
The peak or plateau that  {\sc flock}'s deterministic algorithm finds is entirely dependent on the initial allocation of 
individuals, as it has no way to visit areas of the surface with smaller likelihood values.
If the number of initial allocations does not increase
as $K$ is increased, the starting allocations become diffusely scattered in the space of all
possible partitions making it unlikely that they will converge. This coupled with the fact that there is 
ambiguity in what defines a "best run" when all plateaus are of length zero complicates the interpretation
of results from {\sc flock} if no non-zero plateaus are encountered.

One of the major claimed advantages of {\sc flock} over other methods for genetic clustering
is that, ``The very short processing times for each run of {\sc flock} allows for
comparing partitions from many runs of each $K$'' \citep[][p.~735]{Duc&Tur2012}.
While the authors observed faster processing times (relative to other methods) in their 
comparisons, we did not observe the same advantage in our analysis. This appears to result from two issues. First,
our example data set was an order of magnitude larger than the largest data set analyzed in \citet{Duc&Tur2012}.  Second, 
in the comparison by \citet{Duc&Tur2012}, 
{\sc flock} was set to perform 50 runs and 20 iterations for each $K$ from 1 to four more than 
the true value of $K$, and computing times for that analysis were compared to times for ten iterations of {\sc structure} using 
a 50,000 sweep burn-in followed 
by a 200,000 sweep sample from the posterior of $K$ from 1 to three more than the true 
value of $K$. 
To us it seemed reasonable to compare, for each $K$, one iteration of {\sc structure} (50,000 burn in and 200,000 sample collection sweeps)
to one iteration of {\sc flock} (50 different starting allocations that are reallocated 20 times) 
as both are used to provide a single estimate of the partition of individuals into clusters. One could
argue that it would be more appropriate to compare the running times of the programs according to the number
of opportunities that each individual's cluster membership has for random initialization and reallocation.  Under 
such a perspective, {\sc flock}'s 50 different starting allocations and 20 reallocations should be compared to
50 separate runs of structure, each of length 20 sweeps, which would surely give {\sc structure}
an apparent advantage in speed.  Regardless, while it may hold that {\sc flock}
is able to process small datasets quickly enough to compete with {\sc structure},
the rapid increase in our ability to economically
genotype a large number of individuals for a great many number of loci, may render this advantage less relevant to many 
study designs in the future.

For those that do decide to use {\sc flock} as an alternative to {\sc structure}, the adhoc rules 
to estimate K may be overly stringent. While there has been ample 
discourse concering the choice of the optimal K in {\sc structure} \citep{Evannoetal2005, 
Pritchardetal2000, Wap&Gag2006} much less effort has been put forth examining how the number of individuals
and the number of starting allocations can affect not only the convergence of different runs, but also the 
inference of $K$. The authors developed estimation rules based on simulated data 
sets with a modest number of individuals and different migration regimes and rates, number of loci, and K. 
By relying on identical allocations (mean LLOD values) as a means of supporting 
a particular $K$,  similar but not idential 
allocations will provide no support to a given $K$. We highly advise, if using {\sc flock}, to
run multiple small batches and using the programs {\sc clumpp} \citep{Jak&Ros2007} and
{\sc district} \citep{Rosenberg2004} to visually inspect the 'best run' results.

The development of software to analyze genetic data is ever increasing. In the face of a growing
number of programs, end users must wade through a seemingly countless number of programs each with
their own assumptions, strengths, and weaknesses. Lacking a recognition of these properties
can result in the ineffectual use of both time and resources.
By providing a different prespective on the algorithm
implemented in {\sc flock} we hope to increase understanding of its applicability to a different datasets
while decreasing end user confusion and misuse.
